Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-08-30 Manuscript.Rmd
Added info to intro and more methods

---
title: "Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach"
author: "Claudio Toro-Serey & Joseph T. McGuire"
#date: "7/2/2018"
output:
  pdf_document: default
  html_document: default
---

@@

  A persistent issue in analyzing mPFC in individuals is the various factors inducing signal dropout in its more ventral segments (Logothetis, 2008). As such, there is value in examining all possible sources of covariation that compose DMN, instead of estimating topologies one seed at a time. Connectome-based analyses of resting state functional connectivity (rsFC) have been fruitful in characterizing individualized functional topologies that match task-induced activity (Gordon et al., 2017; Laumann et al., 2015; Tobyne et al., 2018). A popular approach to analyze these connectomes is to rely on graph theoretic methods, which provide an organic avenue to understand brain network dynamics (Bassett, Zurn, & Gold, 2018). From these, community detection algorithms have been successfully utilized to section brain networks into cohesive substructures (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018). Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network (Fortunato & Hric, 2016). Among the algorithms used to estimate these communities, modularity has been widely effective in subsectioning brain networks into multiple groups (Garcia et al., 2018). However, in attempting to refine the DMN topology at the individual level, dividing these regions into what belongs and does not belong to DMN makes dissection through spectral partitioning a more valuable option. In short, spectral partitioning performs eigenvalue decompositon on an adjusted form of a correlation matrix (Laplacian matrix). The resulting eigenvector corresponding to the second-to-last eigenvalue (i.e. the Fiedler Vector) provides values to divide the network into a positive and a negative community (Fiedler, 1975). This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

@@

$$\begin{aligned}
L = I - D^{-\frac{1}{2}}AD^{-\frac{1}{2}}
\end{aligned}$$

Where I is the identity matrix, D is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and A is the correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler Vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected (Fiedler, 1975). Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the one DMN-specific area from the meta-analysis. Importantly, given the high density of these networks (**Add to results**), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016).

In order to evaluate the validity of the resulting partitions, we also estimated these communities using the more traditional approach of modularity maximization (using the algorithm from Clauset et al., XXXX). This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combination of vertices, and selects the clustering that maximizes the network's modularity. Thus, unlike SP, modularity can subdivide a network into many functional groups. If the partitions from the bounded and unbounded methods show high level of agreement, we can gain confidence in that we are indeed fractioning the ROIs into those belonging to DMN and not.

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

@@

  A persistent issue in analyzing mPFC in individuals is the various factors inducing signal dropout in its more ventral segments (Logothetis, 2008). As such, there is value in examining all possible sources of covariation that compose DMN, instead of estimating topologies one seed at a time. Connectome-based analyses of resting state functional connectivity (rsFC) have been fruitful in characterizing individualized functional topologies that match task-induced activity (Gordon et al., 2017; Laumann et al., 2015; Tobyne et al., 2018). A popular approach to analyze these connectomes is to rely on graph theoretic methods, which provide an organic avenue to understand brain network dynamics (Bassett, Zurn, & Gold, 2018). From these, community detection algorithms have been successfully utilized to section brain networks into cohesive substructures (Garcia, Ashourvan, Muldoon, Vettel, & Bassett, 2018). Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network (Fortunato & Hric, 2016). Among the algorithms used to estimate these communities, modularity has been widely effective in subsectioning brain networks into multiple groups (Garcia et al., 2018). However, in attempting to refine the DMN topology at the individual level, dividing these regions into what belongs and does not belong to DMN makes dissection through spectral partitioning (SP) a more valuable option. In short, spectral partitioning performs eigenvalue decompositon on an adjusted form of a correlation matrix (Laplacian matrix). The resulting eigenvector corresponding to the second-to-last eigenvalue (i.e. the Fiedler vector) provides values to divide the network into a positive and a negative community (Fiedler, 1975). SP has recently been used to characterize the posterior-anterior functional gradient of the insula using rsfMRI data (BLAH.) This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

@@

$$\begin{aligned}
L = I - D^{-\frac{1}{2}}WD^{-\frac{1}{2}}
\end{aligned}$$

Where I is the identity matrix, D is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and W is the correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler Vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected (Fiedler, 1975). Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (**Add to results**), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016).

In order to evaluate the validity of the resulting partitions, we also estimated these communities using the more traditional approach of modularity maximization (using the algorithm from Clauset et al., XXXX). This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combination of vertices, and selects the clustering that maximizes the network's modularity. Thus, unlike SP, modularity can fractionate a network into many functional groups. If the partitions from the bounded and unbounded methods show high level of agreement, we can gain confidence in that we are indeed fractioning the ROIs into those belonging to DMN and not.
